---
title: "Dataset preparation"
output: html_document
date: "2023-02-26"
bibliography: references.bib
---

## Data 

This dataset was extracted from  [Tidytuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-10/readme.md) and described in detail in @dataset-article.

The species code data was extracted from the official webpage of [Feederwatch project ](https://feederwatch.org/explore/raw-dataset-requests/) (_accessed `r file.info("data/data.csv")$mtime`_)

#### Variables of interest

**_id**
- `loc_id`
- `project_period_id`
- `obs_id` - should be unique
- `sub_id`

**geographical data**
- `subnational1_code` - Canada provinces
- `lat`
- `long`

**sps data**
- `species_code`

**data quality/description**
- `valid`
- `review`
- `effort_hrs_atleast`
- `day*_*`
- `how_many`
- `Data_Entry_Method`
- `date`
- `entry_technique`

**climate**
- `snow_dep_atleast`

**location**
- Habitat (`hab_*`)
- population_atleast (`*_atleast`)
- count_area_size_sq_m_atleast (`*_atleast`)
- housing_density
- minimum number of tress and other vegetation (`*_atleast`)
- `yard_`

**feeders**
- `nearby_feeders`
- When are the birds feed? (`fed_*`) - (connected with `Date`)
- Number of each type of feeders (`numfeeders-*`) - (connected with `fed_*`)
- squirrels

**Presence**
- cats
- dogs
- humans



```{r download-data}

library(readr)
library(here)

feederwatch <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-10/PFW_2021_public.csv')

site_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-10/PFW_count_site_data_public_2021.csv')

species_code_data <- readr::read_csv(here('data', 'PFW-species-translation-table.csv'))


```

## Data wrangling and saving the file


```{r data-analysis}

library(dplyr)
library(tidyr)
library(stringr)

# issue: Joining, by = "loc_id"
data <-  dplyr::left_join(feederwatch,
                          site_data,
                          by =c('loc_id' = 'loc_id',
                                'PROJ_PERIOD_ID' = 'proj_period_id')) |> 
         dplyr::left_join(species_code_data) |> # Joining, by = "species_code"
         dplyr::filter(str_detect(subnational1_code,
                                  "CA-")) |> # Remove US data
         tidyr::unite("date", c('Month', 'Day', 'Year'),
                      sep = "/") |> # Combine date information
         dplyr::mutate(date = as.Date(date,
                                      '%m/%d/%Y'))   # Change data format
         

write.csv(data, here('data', 'data.csv'))


```

## Data summary

```{r summary}

summary(data)

```


```{r sps-number}

# I have too many species!
unique(data$species_code)

```



## Data exploration

#### Is there more than one observation by location?

```{r}

## Point overlapping
data |> 
  dplyr::count(loc_id,
               species_code,
               latitude,
               longitude) |> 
  filter(species_code == 'norcar') |> 
  arrange(desc(n))


```

## What provinces/territories the species have been seen?

```{r}

library(dplyr)

data |> 
  group_by(species_code) |> 
  summarize(n = n(), 
            provs = paste(subnational1_code, collapse = ", "))

```


## References






